
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(engine.opts = list(bash = "-l"))
```


```{css, echo=FALSE}
h1, h2 {
  text-align: center;
  color: black;
  font-weight: bold;
}
```

## Project materials

- download from Github (https://github.com/yanghong-bio/BBI03X-Project)

## Download metadata (1)

- Rstudio
- install package 'BiocManager' (only need once)

```{r, message=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install()
```

- install other packages (only need once)
```{r, message=FALSE}
BiocManager::install(c("DESeq2", "pheatmap","tidyverse","xlsx","readxl","gplots","ggbiplot",
                       "piano","venn","clusterProfiler","GEOquery"), ask = F, update = F)
```

## Download metadata (2)

- Rstudio
- load packages (need each time you run your code when you come back from )

```{r, message=FALSE}
library(DESeq2)
library(pheatmap)
library(tidyverse)
library(xlsx)
library(readxl)
library(gplots)
library(ggbiplot)
library(piano)
library(venn)
library(clusterProfiler)
library(GEOquery)
library(openxlsx)
library(GEOquery)
```

## Download metadata (2)

- Rstudio
- download datasets from GEO database

```{r, message=FALSE}
GSE135251 <- getGEO('GSE135251',GSEMatrix=TRUE)
GSE135251 <- GSE135251[[1]]
sampleInfo <- pData(GSE135251)
write.xlsx(sampleInfo, file = "./data/DGSE135251_metadata.xlsx")
sampleInfo[1:4,1:3]
```

## Load count data (pre-processed by Authors)

- Rstudio

```{bash, engine.opts='-l', message=FALSE}
# enter data folder 
cd data

# download files
wget https://www.ncbi.nlm.nih.gov/geo/download/\?acc=GSE135251\&format=file -O GSE135251_RAW.tar

#create a new folder for count files
mkdir countfiles

# unzip raw data file and put all the files that it has into new folder 
tar xzf GSE135251_RAW.tar -C countfiles

# enter the folder and unzip all files
cd countfiles
gunzip *.gz
```

## Save count data


```{r}
result_dir = './data/countfiles/'

count_tr = ''
for(i in list.files(result_dir,recursive=F)){
  temp = read.csv(paste0(result_dir, i),sep='\t',stringsAsFactors = F)
  id = gsub("_.*", "",i)
  colnames(temp) = c("ID","Count")
  temp$Sample = id
  count_tr = rbind(count_tr,temp)
}

count_tr = count_tr[-1,]
count_tr_reshape = count_tr %>% spread(Sample, Count)
write.xlsx(count_tr_reshape, file = './data/DGSE135251_sample2count_str.xlsx')
```

